# Reproducible Research: Peer Assessment 1
---
**Title: "Reproductible Research - First Assignment"**  
author: *"Iosif Tanasescu" !*  
date: "Sunday, August 10, 2014"  
output: html_document  
---
```{r,echo=FALSE}
library(knitr)
library(xtable)
```

<!-- 
    Here we are ensuring the echo is always ON 
    This code is hidden in HTML output -->
    
```{r setoptions, echo=FALSE}
opts_chunk$set(echo =  TRUE, results ="asis")
```

## Loading and preprocessing the data  

First we read from the zip file the activity.csv file which contains our data.  
Then we convert the date colum from string to date type  
<!--Finally discard the NA cases-->

```{r}
df <- read.csv(unz("activity.zip","activity.csv"), header =TRUE, sep=",")
df$date <- as.Date(df$date,"%Y-%m-%d")
#df <-df[complete.cases(df),]
```

## What is mean total number of steps taken per day?

Histogram of summary of steps by day  
```{r}
aa<- as.data.frame(tapply(df$steps, df$date , sum, ns.rm=TRUE))
names(aa) <- c("summary") 
hist2 <- cbind(date=rownames(aa),aa)
hist(hist2$summary, xlab="Summary of steps by day", main="Histogram of summary of steps on a daily basis")
```

Mean and median by day  
```{r}
aa<- as.data.frame(tapply(df$steps, df$date , mean, na.rm=TRUE))
aa <- cbind(aa,                  tapply(df$steps, df$date , median, na.rm=TRUE))
names(aa) <- c("daily_mean","daily_median") 
hist1 <- cbind(date=rownames(aa),aa)
#print(xtable(hist1[2:3]))
print(xtable(hist1[2:3]),type="html")
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
