# Reproducible Research: Peer Assessment 1
---
**Title: "Reproductible Research - First Assignment"**  
author: *"Iosif Tanasescu"*  
date: "Sunday, August 10, 2014"  
output: html_document  
---
<!--
     In order to see this html file paste in your browser the following link
     http://htmlpreview.github.io/?https://raw.githubusercontent.com/iositan/RepData_PeerAssessment1/master/PA1_template.html
-->
```{r libraries}
#clean and loading required libraries
rm(list =ls())
library(knitr)
library(xtable)
```

<!-- 
    Here we are ensuring the echo is always ON 
    This code is hidden in HTML output -->
    
```{r setoptions}
opts_chunk$set(echo =  TRUE, results ="asis")
options(digits =3)
```

## Loading and preprocessing the data  

First we read from the zip file the activity.csv file which contains our data.  
Then we convert the date colum from string to date type  
<!--Finally discard the NA cases-->

```{r read_data}
df <- read.csv(unz("activity.zip","activity.csv"), header =TRUE, sep=",")
df$date <- as.Date(df$date,"%Y-%m-%d")
```

## What is mean total number of steps taken per day?

Histogram of summary of steps by day  
```{r plot_histogram}
aa<- as.data.frame(tapply(df$steps, df$date , sum, ns.rm=TRUE))
names(aa) <- c("summary") 
hist2 <- cbind(date=rownames(aa),aa)
hist(hist2$summary, xlab="Summary of steps by day", main="Histogram of summary of steps on a daily basis")
```

Mean and median by day  
```{r mean_and_median_by_day}
# compute the mean
aa<- as.data.frame(tapply(df$steps, df$date , mean, na.rm=TRUE))
# compute the median
aa <- cbind(aa,  tapply(df$steps, df$date , median, na.rm=TRUE))
# Settings names for the columns
names(aa) <- c("Daily Mean","Daily Median") 
# Building the data frame 
hist1 <- cbind(date=rownames(aa),aa)
#print(xtable(hist1[2:3]))
# printing in the HTML format
print(xtable(hist1[2:3]),type="html")
```

## What is the average daily activity pattern?
```{r mean_by_steps}
# compute the mean by steps 
am <- as.data.frame(tapply(df$steps,df$interval,mean, na.rm=TRUE))
names(am) <- c("average_5")
#aa <- cbind(time=rownames(aa),aa)
#  build our time serie
plot3 <- as.ts(as.matrix(am))
plot.ts(plot3)
```

```{r computetime,results=FALSE}
# find out the maximum and associated step
maxidx <- which.max(plot3)
idx <- names(am[maxidx,1])
maxval <- max(plot3)
```
The **index `r idx` of 5-minute interval** (the `r maxidx` x value), on average across all the days in the dataset, contains the **maximum number `r maxval`** of steps


## Imputing missing values

```{r missing_values}
# Missing Values section ---
# Find how many missing rows are in ther initial 
missing <- sum(is.na(df))
# which is
#missing <- nrow(df) - nrow(df[complete.cases(df),])
```
There are **`r missing` rows** in the initial dataset with data missing

We will fill the missing values as follows:  
  - in the steps column, we will replace the NA with 0
```{r Nas,results=FALSE}
# simply replacement 
dfc <- df ; dfc[is.na(dfc)] <- 0

# a litle bit more sofisticated
#
# prepare the means results
#cam <- am ;cam$interval <- rownames(cam)
#rownames(cam) <- 1:nrow(cam)
#dfc <- df ;
#dfc$steps[is.na(dfc$steps)] <- cam$average_5[cam$interval == dfc$interval]

aac<- as.data.frame(tapply(dfc$steps, dfc$date , sum, ns.rm=TRUE))
names(aac) <- c("summary") 
hist2c <- cbind(date=rownames(aac),aac)
hist(hist2c$summary, xlab="Altered Summary of steps by day", main="Histogram of altered summary of steps on a daily basis",col="blue")

aac<- as.data.frame(tapply(dfc$steps, dfc$date , mean, na.rm=TRUE))
# compute the median
aac <- cbind(aac,  tapply(dfc$steps, dfc$date , median, na.rm=TRUE))
# Settings names for the columns
names(aac) <- c("Daily Mean","Daily Median") 
# Building the data frame 
hist1c <- cbind(date=rownames(aac),aac)
# printing in the HTML format
```

```{r t,comments=FALSE}
print(xtable(hist1c[2:3]),type="html")
``` 

## Are there differences in activity patterns between weekdays and weekends? ##
```{r differences}
dfc$wd <- weekdays(dfc$date)
drw <- lapply(dfc$wd, function(x) ifelse(x =="Saturday" || x == "Sunday" , "weekend","weekday"))
dfwa <- c(do.call("cbind",drw))
dfc <-cbind(dfc,dfwa)
dfc$dfwa <- as.factor(dfc$dfwa)

library(lattice)
xyplot(dfc$steps ~ dfc$interval  | dfc$dfwa ,layout = c(1,2), type="l", xlab = "Interval", ylab = "Number of steps")
```

